---
title: "Power simulation for smile"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Load libraries and data
```{r setup, include= F}
# load libraries
library(tidyverse)
library(simr)
library(lmerTest)
library(broom.mixed)
library(emmeans)
library(ggplot2)
library(gridExtra)
library(parallel)

# turn off scientific notation
options(scipen = 999)

# set seed
set.seed(1967)
```

```{r}
# load data
DF <- readRDS(here("power simulation", "DF.l.full.rds")) %>%
  
  # focus on conditions being replicated
  filter(image == 'present', 
         condition == "mimicry",
         trial != 'fil.1',
         trial != 'fil.4') %>% 
  
  # refactor trial
  mutate(trial = if_else(trial == "neutr",
                         "a.neutr",
                         "b.happy"),
           trial = factor(trial)) %>% 
  
  # select relevant variables
  select(ResponseId, trial, happiness)
```

# Fit model based on pilot data
```{r}
pilot.hap <-
    lmer(happiness ~ trial +
         (1 | ResponseId),
         data = DF)
```

# Create data simulation function
```{r}
DataSim <- function(n.sub = 100,
                    n.trial = 2, # participants will complete two trials (happy and neutral pose)
                    fixef = pilot.hap@beta,
                    VarCorr = VarCorr(pilot.hap),
                    sigma = sigma(pilot.hap),
                    stimulus, # label for stimuli valence
                    repetition, # label for repetition condition
                    threat){ # label for threat condition
  
  # create empty dataframe
  dat <- matrix(data = NA,
                nrow = n.sub * n.trial, 
                ncol = 0) %>% 
    data.frame()
    
  # create level 2 vector
  dat$sub = rep(x = 1 : n.sub, 
                each = n.trial) %>% 
    as.factor()
  
  # create level 1 vectors 
  dat$trial = rep(x = levels(DF$trial),
                  each = n.trial / 2,
                  times = n.sub) %>% 
    as.factor()
    
  # simulate happiness and anger scores
  dat$hap <-
    # fit lmer
    makeLmer(formula = y ~ trial + (1 | sub),
             fixef = fixef,
             VarCorr = VarCorr,
             sigma = sigma,
             data = dat) %>% 
    
    # use lmer to simulate happiness scores
    doSim()
  
  # add condition labels
  dat$stimulus = stimulus
  dat$repetition = repetition
  dat$threat = threat
  
  return(dat)
  }
```

# Function that simulates a dataframe with the main-effect of pose
```{r}
MainSim <- function(n.sub = 500){
  c.1 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2, 
            fixef = pilot.hap@beta, # simple effect of pose = pilot effect
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap),  # from previous work
            # conditions: positive stimulus, no repetition, no threat
            stimulus = 'pos',
            repetition = 'once',
            threat = 'false') 
  
  c.2 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2, # number of trials = 2
            fixef = pilot.hap@beta, # effect of pose = pilot effect
            VarCorr = VarCorr(pilot.hap), # variance = pilot variance
            sigma = sigma(pilot.hap), # sigma = pilot sigma
            # conditions: negative stimulus, no repetition, no threat
            stimulus = 'neg',
            repetition = 'once',
            threat = 'false')
  
  c.3 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2,
            fixef = pilot.hap@beta, # simple effect of pose = pilot effect
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
            # conditions: positive stimulus, multiple repetitions, no threat
            stimulus = 'pos',
            repetition = 'multi',
            threat = 'false')

  c.4 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2,
            fixef = pilot.hap@beta, # simple effect of pose = pilot effect
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
            # conditions: negative stimulus, multiple repetitions, no threat
            stimulus = 'neg',
            repetition = 'multi',
            threat = 'false')
  
  c.5 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2, 
            fixef = pilot.hap@beta, # simple effect of pose = pilot effect
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap),  # from previous work
            # conditions: positive stimulus, no repetition, no threat
            stimulus = 'pos',
            repetition = 'once',
            threat = 'true') 
  
  c.6 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2, # number of trials = 2
            fixef = pilot.hap@beta, # effect of pose = pilot effect
            VarCorr = VarCorr(pilot.hap), # variance = pilot variance
            sigma = sigma(pilot.hap), # sigma = pilot sigma
            # conditions: negative stimulus, no repetition, no threat
            stimulus = 'neg',
            repetition = 'once',
            threat = 'true')
  
  c.7 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2,
            fixef = pilot.hap@beta, # simple effect of pose = pilot effect
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
            # conditions: positive stimulus, multiple repetitions, no threat
            stimulus = 'pos',
            repetition = 'multi',
            threat = 'true')

  c.8 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2,
            fixef = pilot.hap@beta, # simple effect of pose = pilot effect
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
            # conditions: negative stimulus, multiple repetitions, no threat
            stimulus = 'neg',
            repetition = 'multi',
            threat = 'true')
  
  # rbind
  df.sim <- rbind(c.1, c.2, c.3, c.4,
                  c.5, c.6, c.7, c.8)
  
  # add an identification variable for each subject
  df.sim$sub <- rep(1:(8*n.sub), each = 2)
  
  return(df.sim)
}
```

# Function that simulates a data frame with a single two-way interaction effect
```{r}
TwoWaySim <- function(n.sub = 500){
  c.1 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2, 
            fixef = pilot.hap@beta, # simple effect of pose = pilot effect
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
            # conditions: positive stimulus, no repetition, no threat
            stimulus = 'pos',
            repetition = 'once',
            threat = 'false')
  c.2 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2,
            fixef = c(pilot.hap@beta[1], 
                      0), # simple effect of pose = 0
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
            # conditions: negative stimulus, no repetition, no threat
            stimulus = 'neg',
            repetition = 'once',
            threat = 'false')
  
  c.3 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2,
            fixef = pilot.hap@beta, # simple effect of pose = 0
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
             # conditions: positive stimulus, multiple repetition, no threat
            stimulus = 'pos',
            repetition = 'multi',
            threat = 'false')

  c.4 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2,
            fixef = c(pilot.hap@beta[1],
                      0), # simple effect of pose = 0
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
            # conditions: negative stimulus, multiple repetition, no threat
            stimulus = 'neg',
            repetition = 'multi',
            threat = 'false')

  c.5 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2, 
            fixef = pilot.hap@beta, # simple effect of pose = pilot effect
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
            # conditions: positive stimulus, no repetition, with threat
            stimulus = 'pos',
            repetition = 'once',
            threat = 'true')
  c.6 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2,
            fixef = c(pilot.hap@beta[1], 
                      0), # simple effect of pose = 0
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
            # conditions: negative stimulus, no repetition, with threat
            stimulus = 'neg',
            repetition = 'once',
            threat = 'true')
  
  c.7 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2,
            fixef = pilot.hap@beta, # simple effect of pose = 0
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
             # conditions: positive stimulus, multiple repetition, with threat
            stimulus = 'pos',
            repetition = 'multi',
            threat = 'true')

  c.8 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2,
            fixef = c(pilot.hap@beta[1],
                      0), # simple effect of pose = 0
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
            # conditions: negative stimulus, multiple repetition, with threat
            stimulus = 'neg',
            repetition = 'multi',
            threat = 'true')
  
  # rbind
  df.sim <- rbind(c.1, c.2, c.3, c.4,
                  c.5, c.6, c.7, c.8)
  
  # add an identification variable for each subject
  df.sim$sub <- rep(1:(8*n.sub), each = 2)
  
  return(df.sim)
}
```

#  Function that simulates a data frame with a single three way interaction effect
```{r}
ThreeWaySim <- function(n.sub = 500){
  c.1 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2, 
            fixef = pilot.hap@beta, # simple effect of pose = pilot effect
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
            # conditions: positive stimulus, no repetition, no threat
            stimulus = 'pos',
            repetition = 'once',
            threat = 'false')
  c.2 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2,
            fixef = c(pilot.hap@beta[1], 
                      0), # simple effect of pose = 0
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
            # conditions: negative stimulus, no repetition, no threat
            stimulus = 'neg',
            repetition = 'once',
            threat = 'false')
  
  c.3 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2,
            fixef = pilot.hap@beta, # simple effect of pose = 0
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
             # conditions: positive stimulus, multiple repetition, no threat
            stimulus = 'pos',
            repetition = 'multi',
            threat = 'false')

  c.4 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2,
            fixef = c(pilot.hap@beta[1],
                      0), # simple effect of pose = 0
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
            # conditions: negative stimulus, multiple repetition, no threat
            stimulus = 'neg',
            repetition = 'multi',
            threat = 'false')

  c.5 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2, 
            fixef =  c(pilot.hap@beta[1], 
                      0), # simple effect of pose = 0
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
            # conditions: positive stimulus, no repetition, with threat
            stimulus = 'pos',
            repetition = 'once',
            threat = 'true')
  c.6 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2,
            fixef = c(pilot.hap@beta[1], 
                      0), # simple effect of pose = 0
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
            # conditions: negative stimulus, no repetition, with threat
            stimulus = 'neg',
            repetition = 'once',
            threat = 'true')
  
  c.7 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2,
            fixef =  c(pilot.hap@beta[1], 
                      0), # simple effect of pose = 0
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
             # conditions: positive stimulus, multiple repetition, with threat
            stimulus = 'pos',
            repetition = 'multi',
            threat = 'true')

  c.8 <-
    DataSim(n.sub = n.sub, 
            n.trial = 2,
            fixef = c(pilot.hap@beta[1],
                      0), # simple effect of pose = 0
            VarCorr = VarCorr(pilot.hap), # from previous work
            sigma = sigma(pilot.hap), # from previous work
            # conditions: negative stimulus, multiple repetition, with threat
            stimulus = 'neg',
            repetition = 'multi',
            threat = 'true')
  
  # rbind
  df.sim <- rbind(c.1, c.2, c.3, c.4,
                  c.5, c.6, c.7, c.8)
  
  # add an identification variable for each subject
  df.sim$sub <- rep(1:(8*n.sub), each = 2)  
  return(df.sim)
}
```

# Power simulation function that returns whether the global and simple effects of interest were significant
```{r}
powSim <- function (df.sim) {
  
  # fit model
  sim.m <-
    lmer(hap ~ trial * stimulus * repetition * threat + 
           (1 | sub),
         data = df.sim)
  
  # extract global effects of interest and determine significance
  global_eff <- sim.m %>% 
    tidy(effects = 'fixed') %>%
    mutate(n = length(unique(df.sim$sub)),
           test = "global",
           condition = term,
           sig = ifelse(p.value < 0.05,
                        1, 0)) %>%
    filter(condition %in% 
             c(
               # main effects
               "trialb.happy", "stimuluspos", "repetitiononce", "threatfalse",
               # two-way interactions
               "trialb.happy:stimuluspos", "trialb.happy:threatfalse", 
               #three-way interaction
               "trialb.happy:stimuluspos:threatfalse"
               )) %>%
    select(n, test, condition, sig)
  
  # test simple effects
  EMM <- emmeans(sim.m,
              specs=pairwise~trial|stimulus:repetition:threat,
              level=0.95,
              adjust="tukey",
              pbkrtest.limit = 8000)
  
  # evaluate simple effects
  EMM_summ <- summary(EMM, infer=TRUE)$contrasts %>%
  mutate(condition = paste0(stimulus, repetition, threat))
  
  # extract simple effects of interest and determine significance
  simple_eff <- EMM_summ %>%
    mutate(n = length(unique(df.sim$sub)),
           test = "simple",
           sig = ifelse(p.value < 0.05,
                        1, 0)) %>%
    select(n, test, condition, sig)
  
  # bind the global and simple effects
  power_table <- rbind(global_eff, simple_eff)
  
  #return the effects power table
  return(power_table)
}
```


# Power plot for the Main Effect model
```{r}
# track simulation time
start_time <- Sys.time()

# simulate power for the main effect model
# expected result: all simple effects are significant

iterations <- 1000 # number of iterations for the power analysis
n_list <- c(25, 50, 100, 150, 200) # sample size list for power analysis


#Declare the number of cores
numCores <- 20 
#Create cluster
cl <- makeCluster(numCores)
#Export any variables you might need in the computation
##Tips: export all variables and functions you will use in the parLapply function, as well as the arguments (if using large data frames, remove any unnecessary columns)
##exporting will take some time and increases memory use, but the parallel processing will run much faster. 
clusterExport(cl, varlist = c("iterations", "n_list", "DF", "pilot.hap", "powSim", "MainSim", "TwoWaySim", "ThreeWaySim", "DataSim"))
#Load the libraries required for the computation in all cores
clusterEvalQ(cl, { library('tidyverse')
library('simr')
library('lmerTest')
library('broom.mixed')
library('emmeans') })


# apply the power simulation function across the sample size list, according to the number of iterations, for the Main Effect model 
pwr.sim.main <- 
      parLapply(cl, n_list, function(n){
        replicate(iterations, 
            powSim(MainSim(n.sub = n)),
            simplify = F) %>% 
        # bind output
         bind_rows()
        }) %>%
  bind_rows

# track simulation time
print(execution_time <- Sys.time() - start_time)

# Optional: save or retrieve data from power simulation
# saveRDS(pwr.sim.main, "pwr_sim_main.Rds")  
# pwr.sim.main <- readRDS("pwr_sim_main.Rds")

# filter effects of interest and calculate power
main_df <- pwr.sim.main %>% 
  filter(condition %in% c(
  "posoncefalse",
  "negoncefalse",
  "posmultifalse",
  "negmultifalse",
  "posoncetrue",
  "negoncetrue",
  "posmultitrue",
  "negmultitrue")) %>%
  group_by(n, test, condition) %>%
  # power = proportion of significant tests
  summarise(power = sum(sig) / n()) %>%
  mutate(condition = case_when(
    condition == "posoncefalse" ~ "Positive stimulus / No repetition / No threat",
    condition == "negoncefalse" ~ "Negative stimulus / No repetition / No threat",
   condition == "posmultifalse" ~ "Positive stimulus / Multiple repetitions / No threat",
   condition == "negmultifalse" ~ "Negative stimulus / Multiple repetitions / No threat",
   condition == "posoncetrue" ~ "Positive stimulus / No repetition / With threat",
    condition == "negoncetrue" ~ "Negative stimulus / No repetition / With threat",
   condition == "posmultitrue" ~ "Positive stimulus / Multiple repetitions / With threat",
   condition == "negmultitrue" ~ "Negative stimulus / Multiple repetitions / With threat")) %>%
  rename(Condition = condition)

# plot the power curve for the effects of interest
main_plot <- ggplot(main_df, aes(x = n, y = power, color = Condition, group = Condition)) +
  geom_line(size = 1, alpha = 0.5) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.95, linetype = "dotted", color = "black") +
  coord_cartesian(ylim = c(0, 1)) +
  theme_minimal() +
  labs(
    title = "a) Main Effect Model",   # Plot title
    x = "Total Sample Size",                # X-axis label
    y = "Power"
    )
print(main_plot)
```

# Power plot for the single two-way interaction effect model
```{r}
# expected result: simple effects are significant when feedback is positive, but not when feedback is negative

# apply the power simulation function across the sample size list, according to the number of iterations, for the full moderation interaction effect model
pwr.sim.twoway <- parLapply(cl, n_list, function(n){
        replicate(iterations, 
            powSim(TwoWaySim(n.sub = n)),
            simplify = F) %>% 
        # bind output
         bind_rows()
        }) %>%
  bind_rows()

# Optional: save or retrieve data from power simulation
# saveRDS(pwr.sim.twoway, "pwr_sim_twoway.Rds")  
# pwr.sim.twoway <- readRDS("pwr_sim_twoway.Rds")

# filter effects of interest and calculate power
twoway_df <- pwr.sim.twoway %>% 
  filter(condition %in% c(
  "posoncefalse",
  "negoncefalse",
  "posmultifalse",
  "negmultifalse",
  "posoncetrue",
  "negoncetrue",
  "posmultitrue",
  "negmultitrue")) %>%
  group_by(n, test, condition) %>%
  # power = proportion of significant tests
  summarise(power = sum(sig) / n()) %>%
  mutate(condition = case_when(
    condition == "posoncefalse" ~ "Positive stimulus / No repetition / No threat",
    condition == "negoncefalse" ~ "Negative stimulus / No repetition / No threat",
   condition == "posmultifalse" ~ "Positive stimulus / Multiple repetitions / No threat",
   condition == "negmultifalse" ~ "Negative stimulus / Multiple repetitions / No threat",
   condition == "posoncetrue" ~ "Positive stimulus / No repetition / With threat",
    condition == "negoncetrue" ~ "Negative stimulus / No repetition / With threat",
   condition == "posmultitrue" ~ "Positive stimulus / Multiple repetitions / With threat",
   condition == "negmultitrue" ~ "Negative stimulus / Multiple repetitions / With threat"))  %>%
  rename(Condition = condition)

# plot the power curve for the effects of interest
twoway_plot <- ggplot(twoway_df, aes(x = n, y = power, color = Condition, group = Condition)) +
  geom_line(size = 1, alpha = 0.5) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.95, linetype = "dotted", color = "black") +
  coord_cartesian(ylim = c(0, 1)) +
  theme_minimal() +
  labs(
    title = "b) Two-way Interaction Model",   # Plot title
    x = "Total Sample Size",                # X-axis label
    y = "Power"
    ) 
print(twoway_plot)
```

# Power plot for the single Three-way Interaction Effect Model
```{r}
# expected result: all simple effects significant, but power when feedback is positive is greater than when feedback is negative

# apply the power simulation function across the sample size list, according to the number of iterations, for the partial moderation interaction effect model
pwr.sim.threeway <- parLapply(cl, n_list, function(n){
        replicate(iterations, 
            powSim(ThreeWaySim(n.sub = n)),
            simplify = F) %>% 
        # bind output
         bind_rows()
        }) %>%
  bind_rows()

stopCluster(cl)

# Optional: save or retrieve data from power simulation
# saveRDS(pwr.sim.threeway, "pwr_sim_threeway.Rds")  
# pwr.sim.threeway <- readRDS("pwr_sim_threeway.Rds")

# filter effects of interest and calculate power
threeway_df <- pwr.sim.threeway %>% 
  group_by(n, test, condition) %>%
  filter(condition %in% c(
  "posoncefalse",
  "negoncefalse",
  "posmultifalse",
  "negmultifalse",
  "posoncetrue",
  "negoncetrue",
  "posmultitrue",
  "negmultitrue")) %>%
  group_by(n, test, condition) %>%
  # power = proportion of significant tests
  summarise(power = sum(sig) / n()) %>%
  mutate(condition = case_when(
    condition == "posoncefalse" ~ "Positive stimulus / No repetition / No threat",
    condition == "negoncefalse" ~ "Negative stimulus / No repetition / No threat",
   condition == "posmultifalse" ~ "Positive stimulus / Multiple repetitions / No threat",
   condition == "negmultifalse" ~ "Negative stimulus / Multiple repetitions / No threat",
   condition == "posoncetrue" ~ "Positive stimulus / No repetition / With threat",
    condition == "negoncetrue" ~ "Negative stimulus / No repetition / With threat",
   condition == "posmultitrue" ~ "Positive stimulus / Multiple repetitions / With threat",
   condition == "negmultitrue" ~ "Negative stimulus / Multiple repetitions / With threat"))  %>%
  rename(Condition = condition)

# plot the power curve for the effects of interest
threeway_plot <- ggplot(threeway_df, aes(x = n, y = power, color = Condition, group = Condition)) +
  geom_line(size = 1, alpha = 0.5) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.95, linetype = "dotted", color = "black") +
  coord_cartesian(ylim = c(0, 1)) +
  theme_minimal() +
  labs(
    title = "c) Three-way Interaction Model",   # Plot title
    x = "Total Sample Size",                # X-axis label
    y = "Power"
    ) 
print(threeway_plot)

# record simulation end time and print total execution time
# end_time <- Sys.time()
# print(execution_time <- end_time - start_time)

```

# Combine plots in a single figure
```{r}
# function to extract legend box from the plot
get_legend <- function(myplot) {
  tmp <- ggplotGrob(myplot + theme(legend.position = "right")) # transform into a graphical object and determine the legend position
  legend <- tmp$grobs[which(sapply(tmp$grobs, function(x) x$name) == "guide-box")] # extract the legend grob
  return(legend[[1]])
}

# get the legend from the main plot
shared_legend <- get_legend(main_plot)

# remove legend from individual plots
main_plot <- main_plot + theme(legend.position = "none")
twoway_plot <- twoway_plot + theme(legend.position = "none")
threeway_plot <- threeway_plot + theme(legend.position = "none")

# combine plots and include the shared legend
combined_plot <- grid.arrange(main_plot,
          twoway_plot,
          threeway_plot,
          shared_legend,
          ncol = 2,
          top = "Power curves from the Monte Carlo simulations")
```
